---
title: 'Lab Exercise #1'
output: pdf_document
date: "2024-03-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Warpbreak dataset

```{r}
data("warpbreaks")

```

1. Find out, in a single command, which columns of warpbreaks are either numeric or
integer. What are the data types of each column?

#The breaks  are numeric.
#The wool  are factor.
#The tension are factor.
```{r}
str(warpbreaks)
```


2. How many observations does it have?

54 Observations
```{r}
observations <- nrow(warpbreaks)
observations
```

3. Is numeric a natural data type for the columns which are stored as such? Convert to
integer when necessary.
Ans. Yes, numeric is a natural data type for columns

```{r}
warpbreaks$wool <- as.integer(warpbreaks$wool)
warpbreaks$tension <- as.integer(warpbreaks$tension)


str(warpbreaks)
#View(warpbreaks)
```
4. Error messages in R sometimes report the underlying type of an object rather than the
user-level class. Derive from the following code and error message what the
underlying type. Explain what is the error all about. Do not just copy the error
message that was displayed.

B. Load the exampleFile.txt

1. Read the complete file using readLines.
```{r}
lines<- readLines("exampleFile.txt")
lines
```

2.Separate the vector of lines into a vector containing comments and a vector
containing the data. Hint: use grepl.

```{r}

# Separate the lines into comments and data
comments <- lines[grepl("//", lines)]
comments
data <- lines[!grepl("//", lines)]
data
```

3. Extract the date from the first comment line and display on the screen “It was created
data.”

```{r}
date_line <- lines[grepl("^//", lines)]
date_str <- gsub("^//|Survey data. Created : ", "", date_line[1])

print(paste("It was created", date_str))


```

4. Read the data into a matrix as follows.
a. Split the character vectors in the vector containing data lines by semicolon (;)
using strsplit.

```{r}
list <- strsplit(data, split = ";")
list
```

b. Find the maximum number of fields retrieved by split. Append rows that are
shorter with NA's.
```{r}

max_fields <- max(sapply(list, length))
max_fields
cat("The maximum number of fields retrieved by split are: ",max_fields)

splitList<-lapply(list,function(row) c(row, rep(NA,max_fields-length(row))))
splitList
```

c. Use unlist and matrix to transform the data to row-column format.

```{r}
splitList_unlist<- unlist(splitList)
splitList_unlist
data_matrix <- matrix(splitList_unlist, nrow = length(data),byrow=TRUE)
data_matrix
```

d. From comment lines 2-4, extract the names of the fields. Set these as
colnames for the matrix you just created.

```{r}

field_names <- c("Gender", "Age", "Weight")

colnames(data_matrix) <- field_names
data_matrix


```

